---
title: "Indicators for Pandemic Food Security"
author: "Alejandra Muñoz-Rivera"
date: ""
output:
  html_document: default
---

## Objectives 
1.	Determine whether adult food security status associated with citizenship status 
2.	Determine whether child food security status associated with citizenship status 

## Load packages and open libraries 
```{r, echo=TRUE, message = F}
pacman::p_load(odds.n.ends, blorr, lmtest, car,  broom, tidyverse, jtools, dplyr) 
pacman::p_load(readr, nnet, MASS, funModeling, brant)

# tidyverse:data management and visualization
# odds.n.ends: computing sensitivity and specificity, plotting predictive probabilities and ROC curves for logistic regression
# blorr: model fits of logistic regression
# lmtest: likelihood ratio test
# car: compare coefficients
# broom: get cooks d

# readr: for read txt file
# tidyverse: for various packages
# nnet: Multinomial logistic regresison
# MASS: Ordinal logistic regression
# funModeling: for describe function
# brant: for testing parallel regression assumption
```

## Import csv from github
```{r, echo=TRUE}
FS <- read_csv("/Users/alemunoz/Desktop/BrownSchool/Fall22/ADA/FS_data.csv")
```



```{r, echo=TRUE}
#select final interview outcome (HUFINAL), family income (HEFAMINC), age (PRTAGE), race/ethnicity (PTDTRACE), employment status (PEMLR), household size (HRNUMHOU), adult food security status (HRFS12M1), children's food security status (HRFS12MC) and citizenship (PRCITSHP) status. We're throwing in state (GESTFIPS) for fun. 

#Need to figure out how to drop -1 in age & HRNUMHOU (continuous variables). 
fs2 <- FS %>%
  select(PTDTRACE,HEFAMINC, PRCITSHP,HRFS12M1, HRFS12MC) %>%
    mutate(Race = recode_factor(.x = PTDTRACE, 
                                  `-1` = NA_character_,
                                  `1` = "White",
                                  `2` = "Black",
                                  `3` = "AI/AN",
                                  `4` = "Asian",
                                  `5` = "Hawaiian/Pacific Island", 
                                  `6` = "Black",
                                  `7` = "Asian",
                                  `8` = "Hispanic only",
                                  `9` = "Black",
                                  `10` = "Black",
                                  `11` = "Black",
                                  `12` = "Black",
                                  `13` = "AI/AN",
                                  `14` = "AI/AN",
                                  `15` = "Asian",
                                  `16` = "3 or more races",
                                  `17` = "3 or more races",
                                  `18` = "3 or more races",
                                  `19` = "3 or more races",
                                  `20` = "3 or more races",
                                  `21` = "3 or more races",
                                  `22` = "3 or more races",
                                  `23` = "3 or more races",
                                  `24` = "3 or more races",
                                  `25` = "3 or more races",
                                  `26` = "3 or more races"))%>%
    mutate(Inc = recode_factor(.x = HEFAMINC, 
                                    `-1` = NA_character_,
                                    `1` = "<$20,000",
                                    `2` = "<$20,000",
                                    `3` = "<$20,000",
                                    `4` = "<$20,000",
                                    `5` = "<$20,000",
                                    `6` = "<$20,000",
                                    `7` = "$20,000-$39,999",
                                    `8` = "$20,000-$39,999",
                                    `9` = "$20,000-$39,999",
                                    `10` = "$20,000-$39,999",
                                    `11` = "$40,000-$59,999",
                                    `12` = "$40,000-$59,999",
                                    `13` = "$60,000-$74,999",
                                    `14` = "$75,000-$99,999",
                                    `15` = "$100,000-$149,999",
                                    `16` = ">$150,000"))%>%
    mutate(Cit5= recode_factor(.x = PRCITSHP, 
                                    `-1` = NA_character_,
                                    `1` = "Native, born in U.S.",
                                    `2` = "Native, born in U.S. Island Areas",
                                    `3` = "Native, born abroad of Amer. parent(s)", 
                                    `4` = "Foregin born, naturalized U.S. citzen", 
                                    `5` = "Foregin born, not a U.S. citizen"))%>%
    mutate(Cit3 = recode_factor(.x = PRCITSHP, 
                                    `-1` = NA_character_,
                                    `1` = "Native born",
                                    `2` = "Native born",
                                    `3` = "Native born", 
                                    `4` = "Foregin born, naturalized U.S. citzen", 
                                    `5` = "Foregin born, not a U.S. citizen"))%>%   
    mutate(Cit2 = recode_factor(.x = PRCITSHP, 
                                    `-1` = NA_character_,
                                    `1` = "U.S. Citizen",
                                    `2` = "U.S. Citizen",
                                    `3` = "U.S. Citizen", 
                                    `4` = "U.S. Citizen", 
                                    `5` = "Not a U.S. citizen"))%>%   
    mutate(FoodSec = recode_factor(.x = HRFS12M1, 
                                   `-1` = NA_character_,
                                    `1` = "Food Secure High or Marginal Food Security",
                                    `2` = "Low Food Security",
                                    `3` = "Very Low Food Security",
                                    `-9` = NA_character_))%>%
    mutate(KidFood = recode_factor(.x = HRFS12MC, 
                                   `-1` = NA_character_,
                                    `1` = "Food Secure or Marginal Food Security",
                                    `2` = "Low Food Security",
                                    `3` = "Very Low Food Security",
                                    `-9` = NA_character_)) %>%
  drop_na()

```


```{r}
#  mutate(HUFINAL = case_when(HUFINAL > 1  ~ 0,
#                                HUFINAL == 1 ~ 1),
#        HUFINAL = factor(HUFINAL, 0:1, c(NA_character_, "Complete"))) %>%
```

## Check percentage of data excluded
```{r}
nrow(fs2)/nrow(FS) * 100
```
Might be a lot excluded. But not everyone did the food security questionnaire  

## Let's make sure we have enough cases in each exposure/outcome category. 
```{r}
table(fs2$Cit3, fs2$FoodSec, fs2$KidFood)
```
Had to little in each category when citizenship status had 3 or 5 categories. Was recoded into binary (citizen/non-citizen). 

## Run oridinal regression for adult food security 
```{r}
model1 <- polr(FoodSec ~ Cit2, data=fs2, Hess=TRUE) 

summary(model1)

tidy(model1, conf.int=TRUE, exponentiate = TRUE, p.values=TRUE)
```

## Let's adjust for income category and race/ethnicity. 
```{r}
model2 <- polr(FoodSec ~ Cit2 + Inc + Race, data=fs2, Hess=TRUE)
summary(model2)

tidy(model2, conf.int=TRUE, exponentiate = TRUE, p.values=TRUE)
```

The odds of being FS are 1.10 (95% CI 0.97-1.23) times the odds of being non-FS for respondents who are not U.S. citizens. 
#Income
The odds of being FS are 0.79 (95% CI 0.71-0.88) times the odds of being non-FS for respondents who make between $20,000-$39,999. 

The odds of being FS are 0.48 (95% CI 0.43-0.54) times the odds of being non-FS for respondents who make between $40,000-$59,999. 

The odds of being FS are 0.25 (95% CI 0.22-0.28) times the odds of being non-FS for respondents who make between $60,000-$74,999.

The odds of being FS are 0.14 (95% CI 0.12-0.16) times the odds of being non-FS for respondents who make between $75,000-$99,999.

The odds of being FS are 0.07 (95% CI 0.06-0.09) times the odds of being non-FS for respondents who make between $100,000-$149,999.

The odds of being FS are 0.02 (95% CI 0.02-0.03) times the odds of being non-FS for respondents who make more than $150,000.

#Race
The odds of being FS are 1.51 (95% CI 1.38-1.66) times the odds of being non-FS for respondents who are Black. 

The odds of being FS are 1.39 (95% CI 1.11-1.72) times the odds of being non-FS for respondents who are AI/AN. 

The odds of being FS are 1.17 (95% CI 1.00-1.37) times the odds of being non-FS for respondents who are Asian. 

The odds of being FS are 1.09 (95% CI 0.74-1.58) times the odds of being non-FS for respondents who are Hawaiian/Pacific Island.

The odds of being FS are 1.39 (95% CI 0.84-2.19) times the odds of being non-FS for respondents who are Hispanic only.

The odds of being FS are 3.78 (95% CI 2.39-5.81) times the odds of being non-FS for respondents who are 3 or more races.

## Don't have to to test for linearity since non of the variables are continuous.

## Testing the parallel regression/proportional odds assumption using Brant Test for adult food securiyt 
```{r}
brant(model2) #Assumption holds 
```

#Repeat model for kids food security 
```{r}
model3 <- polr(KidFood ~ Cit2 + Inc + Race, data=fs2, Hess=TRUE)
summary(model2)

tidy(model3, conf.int=TRUE, exponentiate = TRUE, p.values=TRUE)
```


## Testing the parallel regression/proportional odds assumption using Brant Test for adult food security
```{r}
brant(model3) #Assumption holds 
```


